---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Load libraries and data file

```{r}
echo = TRUE
library(ggplot2)

activity <- read.csv("activity.csv", header = TRUE)
str(activity)
```

Modify date variable
```{r}
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```

############ What is mean total number of steps taken per day? #################

Calculate the total number of steps taken per day
```{r}
total_step <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
```

Make a histogram of the total number of steps taken each day
```{r}
hist (total_step$steps, main = "Total Number of Steps per Day", xlab = "Steps per day")
```

Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean (total_step$steps)
median (total_step$steps)
```

############### What is the average daily activity pattern? ######################

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
average_steps <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
plot (average_steps$interval, average_steps$steps, type='l', main="Average Number of Steps Taken", xlab="Intervals", ylab="Average Number of Steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
average_steps$interval[which.max(average_steps$steps)]
```

###################### Imputing missing values #############################

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum (is.na(activity)) 
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
mean (activity$steps, na.rm = T)
imputation <- activity
for (i in average_steps$interval) {imputation[imputation$interval == i & is.na(imputation$steps), ]$steps <- average_steps$steps[average_steps$interval == i]}
head(imputation)
```

Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day. 
```{r}
total_imputation <- aggregate(steps ~ date, data = imputation, sum, na.rm = TRUE)
hist (total_imputation$steps, main = "Total Steps per Day - Imputation", xlab = "Steps per Day")
mean (total_imputation$steps)
median (total_imputation$steps)
```

######### Are there differences in activity patterns between weekdays and weekends? #########
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
imputation$weekdays <- weekdays(imputation$date)
imputation$weeks[(imputation$weekdays == "Saturday" | imputation$weekdays == "Sunday")] <- "weekend"
imputation$weeks[!(imputation$weekdays == "Saturday" | imputation$weekdays == "Sunday")] <- "weekdays"
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r}
library(plyr)
new <- ddply(imputation, c("interval", "weeks"), function(x) apply(x[1],2, mean))
head(new)
library(lattice)
xyplot(steps ~ interval | weeks, data = new, type = "l", xlab = "Interval", ylab = "Number of steps", layout = c(1, 2))
```
